Evolution of a conspicuous dynamic visual signal in an introduced chameleon

### 2021-05-05 ms.Rmd
Corrected MS with wald tests

```{r getlibrary, echo = FALSE, results = "hide"}
# load libraries
devtools::install_github("karthik/wesanderson")
pacman::p_load(lme4, lmerTest, ggplot2, lsmeans, emmeans, multcomp, dplyr, wesanderson, patchwork, latex2exp, png, cowplot, magick, jpeg, clubSandwich)

p_value <- function(x){
  if(x <= 0.0001) {tmp = "< 0.0001"}
  if(x <= 0.001 & x >= 0.0001) {tmp ="< 0.001"}
  if(x <= 0.01 & x >= 0.001) {tmp ="< 0.01"}
  if(x >= 0.01) {tmp = round(x, digits =2)}
  return(tmp)
}

@@

```{r Models, echo = FALSE, results = "hide"}
#########################################################
# Social Context: Male-male Competition Displays
#########################################################
#Load data and clean a bit
Display_1 <- read.table("./data/Display_1.csv", 
                          header=TRUE, sep=",", dec=".", strip.white=TRUE)

@@

    # Get the coefficients, and ANOVA statistics for the paper
          coefAIC_display <- fixef(ConspicDisplayAC)[3]
            CI_AC_display <- data.frame(confint(ConspicDisplayAC))
    anovaConspicDisplayAC <- anova(ConspicDisplayAC)

@@

@@

    # Get ANOVA statistics for model for the paper
         anovaConspicDisplayCC <- anova(ConspicDisplayCC)

@@

#########################################################
# Social Context: Courtship Displays
#########################################################

@@

  # Get model stats for paper
       coefAIC_court <- fixef(ConspicCourtshipAC)[3]
         CI_AC_court <- data.frame(confint(ConspicCourtshipAC))
    courship_anovaAC <- anova(ConspicCourtshipAC)

@@

@@

  # Get the ANOVA stats for paper
    courship_anovaCC <- anova(ConspicCourtshipCC)

@@

@@

  # Fit main effects model
    PredatorAC <- lmer(dL ~ Predator + Population + (1|ID), data = Predator) 
    summary(PredatorAC)

@@

@@

    # Get ANOVA statistics for the paper
    anova_predators_interCC <- anova(PredatorCC)

@@

@@

  # Grab contrasts that we need from the robust esatimtes
  p_displ_AC <- as.numeric(result_DisplayBgrd_pop_back[2,8])
   CI_dis_AC <- data.frame(Estimate=robust_DisplayBgrd_pop_back$beta[3],
                           lwr = robust_DisplayBgrd_pop_back$CI_L[3],
                           upr = robust_DisplayBgrd_pop_back$CI_U[3])

@@

@@

@@

@@

Sexual selection is expected to favour more conspicuous dymamic sexual signals in Hawaii under relaxed predation because higher predation should favour cryptic signals whereas greater conspicuousness of sexual signals should be selected in social contexts. In support of this prediction, male display colouration exhibited in social contexts (i.e., male-male contests and courtship) was more conspicuous to a conspecific in Hawaii compared with Kenya (Male-Male Contests: Acromatic contrast ($dl$); $F_{`r anovaConspicDisplayAC$NumDF[2]`,`r anovaConspicDisplayAC$DenDF[2]`}$ = `r anovaConspicDisplayAC[5][2,]`, p = `r p_value(anovaConspicDisplayAC[6][2,])`; Chromatic contrast ($dS$): $F_{`r anovaConspicDisplayCC$NumDF[2]`,`r anovaConspicDisplayCC$DenDF[2]`}$ = `r anovaConspicDisplayCC[5][2,]`, p = `r p_value(anovaConspicDisplayCC[6][2,])`; Courtship displays: Acromatic contrast ($dl$); $F_{`r courship_anovaAC$NumDF[2]`,`r courship_anovaAC$DenDF[2]`}$ = `r courship_anovaAC[5][2,]`, p = `r p_value(courship_anovaAC[6][2,])`; Chromatic contrast ($dS$): $F_{`r courship_anovaCC$NumDF[2]`,`r courship_anovaCC$DenDF[2]`}$ = `r courship_anovaCC[5][2,]`, p = `r p_value(courship_anovaCC[6][2,])`), and this was  true across all body regions measured (See Supplemental Materials, table S1). Kenyan chameleons from populations where predation is high were significantly more crypic compared to introduced Hawaiian counterparts (Acromatic Contrast between Hawaiian and Kenyan Populations, Male-male contests: $\beta$ = `r coefAIC_display`, 95% CI = `r CI_AC_display[5,1]` to  `r CI_AC_display[5,2]`, df = `r summary(ConspicDisplayAC)$coefficients[3, "df"]`, t =  `r summary(ConspicDisplayAC)$coefficients[3, "t value"]`, p = `r p_value(summary(ConspicDisplayAC)$coefficients[3, "Pr(>|t|)"])`, Courtship: $\beta$ = `r coefAIC_court`, 95% CI = `r CI_AC_court[5,1]` to  `r CI_AC_court[5,2]`, df = `r summary(ConspicCourtshipAC)$coefficients[3, "df"]`, t =  `r summary(ConspicCourtshipAC)$coefficients[3, "t value"]`, p = `r p_value(summary(ConspicCourtshipAC)$coefficients[3, "Pr(>|t|)"])`, see figure 1a & b).

As a consequence of signal release, sexual signals are expected to be more easily detected by predators (i.e., snakes and birds) in Hawaii. Differential predation pressure resulting from snake and bird predation varying would also result in populations varying in their conspicuousness to different predators.  In support of these predictions, Hawaiian chameleons were significantly more conspicous to both snake and bird predators (figure 1b & c; Achromatic Contrast: ($dl$); $F_{`r anova_predators$NumDF[2]`,`r anova_predators$DenDF[2]`}$ = `r anova_predators[5][2,]`, p = `r p_value(anova_predators[6][2,])`). Differences in chromatic contrast between populations, however,  depended on the predator (significant predator by population interaction: $F_{`r anova_predators_interCC$NumDF[3]`,`r anova_predators_interCC$DenDF[3]`}$ = `r anova_predators_interCC[5][3,]`, p = `r p_value(anova_predators_interCC[6][3,])`). Interestingly, Kenyan and Hawaiian lizards were equally detectable by snakes, but Kenyan chameleons were significantly more cryptic from birds compared to Hawaiian chamleons (figure 1c; Snakes (Hawaii-Kenya): `r CI_hyp_test[1,2]`, 95% CI = `r CI_hyp_test[1,3]` to `r CI_hyp_test[1,4]`, p = 1; Birds (Hawaii-Kenya): `r CI_hyp_test[2,2]`, 95% CI = `r CI_hyp_test[2,3]` to `r CI_hyp_test[2,4]`, p = `r CI_hyp_test[2,8]`).

Local adaptation of sexual signals would result in Hawaiian chameleons being more conspicous to conspecifics in the Hawaiian environment compared to their native Kenyan environment. While it was not possible to conduct common garden experiments to provide definitive tests of local adaptation, our results support the local adaptation hypothesis given that Hawaiian chameleons were more conspicous to other chamleeons in the Hawaiian environment compared to the same chameleons in a Kenyan environment in both social contexts (figure 2; Contrast between Hawaiian lizard on Kenyan vs Hawaiian on Hawaiin background - Male-male contests: Achromatic ($dl$), $\beta$ = `r CI_dis_AC[1,1]`, 95% CI = `r CI_dis_AC[1,2]` to `r CI_dis_AC[1,3]`, p = `r p_value(p_displ_AC)`, Chromatic ($dS$) $\beta$ = `r CI_Disp_CC[1,1]`, 95% CI = `r CI_Disp_CC[1,2]` to `r CI_Disp_CC[1,3]`, p = `r p_value(p_displ_CC)`; Courtship: Achromatic ($dl$), $\beta$ = `r CI_AC_Courtship[1,1]`, 95% CI = `r CI_AC_Courtship[1,2]` to `r CI_AC_Courtship[1,3]`, p = `r p_value(p_courtship_AC)`, Chromatic ($dS$) $\beta$ = `r CI_CC_Courtship[1,1]`, 95% CI = `r CI_CC_Courtship[1,2]` to `r CI_CC_Courtship[1,3]`, p = `r p_value(p_courtship_CC)`).

>>>>

```{r getlibrary, echo = FALSE, results = "hide"}
# load libraries
devtools::install_github("karthik/wesanderson")
pacman::p_load(lme4, lmerTest, ggplot2, lsmeans, emmeans, multcomp, dplyr, wesanderson, patchwork, latex2exp, png, cowplot, magick, jpeg, clubSandwich)
source("./R/func.R")

@@

```{r Models, echo = FALSE, results = "hide"}
source("./R/func.R")
#########################################################
# Social Context: Male-male Competition Displays
#########################################################
#Load data and clean a bit
Display_1 <- read.table("./data/Display_1.csv", 
                          header=TRUE, sep=",", dec=".", strip.white=TRUE)

@@

@@

     # Wald test
            Population_ConspicDisplayAC <- Wald_test(ConspicDisplayAC, 
                                             constraints = constrain_zero(3),
                                            cluster = Display_1_IDs$ID, vcov = "CR2")

@@

@@

  # Wald test
            Population_ConspicDisplayCC <- Wald_test(ConspicDisplayCC, 
                                             constraints = constrain_zero(3),
                                            cluster = Display_1_IDs$ID, vcov = "CR2")
#########################################################
# Social Context: Courtship Displays
#########################################################

@@

@@

    # Wald test
            Population_ConspicCourtshipAC <- Wald_test(ConspicCourtshipAC, 
                                             constraints = constrain_zero(3),
                                            cluster = Courtship_IDs$ID, vcov = "CR2")

@@

@@

    # Wald test
            Population_ConspicCourtshipCC <- Wald_test(ConspicCourtshipCC, 
                                             constraints = constrain_zero(3),
                                            cluster = Courtship_IDs$ID, vcov = "CR2")

@@

  #Refit main effects so we can test main effects on own.
    PredatorAC <- lmer(dL ~ Predator + Population + (1|ID), data = Predator) 
    summary(PredatorAC)

@@

     # Wald tests
            back_pred_AC_Pred <- Wald_test(PredatorAC_inter, constraints = constrain_zero(4), 
                                   cluster = Predator$ID, vcov = "CR2")

             Predator_AC_Pred <- Wald_test(PredatorAC, constraints = constrain_zero(2), 
                                  cluster = Predator$ID, vcov = "CR2")

           Population_AC_Pred <- Wald_test(PredatorAC, constraints = constrain_zero(3), 
                                  cluster = Predator$ID, vcov = "CR2")

@@

    #Refit main effects so we can test main effects on own.
      PredatorCC_main <- lmer(dS ~ Predator + Population + (1|ID), data = Predator)

     # Wald tests
            back_pred_CC_Pred <- Wald_test(PredatorCC, constraints = constrain_zero(4), 
                                   cluster = Predator$ID, vcov = "CR2")

             Predator_CC_Pred <- Wald_test(PredatorCC_main, constraints = constrain_zero(2), 
                                  cluster = Predator$ID, vcov = "CR2")

           Population_CC_Pred <- Wald_test(PredatorCC_main, constraints = constrain_zero(3), 
                                  cluster = Predator$ID, vcov = "CR2")

@@

  # Add Wald test to test signifiacnce of each factor, background and population using roust variance
  background_AC_Disp <- Wald_test(DispBgrdAC, constraints = constrain_zero(3), 
                                  cluster = DisplayBgrd$ID, vcov = "CR2")

  Population_AC_Disp <- Wald_test(DispBgrdAC, constraints = constrain_zero(2), 
                                  cluster = DisplayBgrd$ID, vcov = "CR2")

@@

  # Grab contrasts that we need from the robust esatimtes
  p_displ_AC <- result_DisplayBgrd_pop_back[2,8]
   CI_dis_AC <- data.frame(Estimate=robust_DisplayBgrd_pop_back$beta[3],
                           lwr = robust_DisplayBgrd_pop_back$CI_L[3],
                           upr = robust_DisplayBgrd_pop_back$CI_U[3])

@@

    # Check significance of each factor
        background_CC_Disp <- Wald_test(DispBgrdCC, constraints = constrain_zero(3), 
                                  cluster = DisplayBgrd$ID, vcov = "CR2")

  Population_CC_Disp <- Wald_test(DispBgrdCC, constraints = constrain_zero(2), 
                                  cluster = DisplayBgrd$ID, vcov = "CR2")

@@

   # Wald tests
           background_AC_Court <- Wald_test(CourtshipAC, constraints = constrain_zero(3), 
                                  cluster = CourtshipBgrd$ID, vcov = "CR2")

           Population_AC_Court <- Wald_test(CourtshipAC, constraints = constrain_zero(2), 
                                  cluster = CourtshipBgrd$ID, vcov = "CR2")

@@

   # Wald tests
           background_CC_Court <- Wald_test(CourtshipCC, constraints = constrain_zero(3), 
                                  cluster = CourtshipBgrd$ID, vcov = "CR2")

           Population_CC_Court <- Wald_test(CourtshipCC, constraints = constrain_zero(2), 
                                  cluster = CourtshipBgrd$ID, vcov = "CR2")

@@

Sexual selection is expected to favour more conspicuous dymamic sexual signals in Hawaii under relaxed predation because higher predation should favour cryptic signals whereas greater conspicuousness of sexual signals should be selected in social contexts. In support of this prediction, male display colouration exhibited in social contexts (i.e., male-male contests and courtship) was more conspicuous to a conspecific in Hawaii compared with Kenya (Male-Male Contests: Acromatic contrast ($dl$); $F_{`r Population_ConspicDisplayAC$df_num`,`r Population_ConspicDisplayAC$df_denom`}$ = `r Population_ConspicDisplayAC$Fstat`, p = `r p_value(Population_ConspicDisplayAC$p_val)`; Chromatic contrast ($dS$): $F_{`r Population_ConspicDisplayCC$df_num`,`r Population_ConspicDisplayCC$df_denom`}$ = `r Population_ConspicDisplayCC$Fstat`, p = `r p_value(Population_ConspicDisplayCC$p_val)`; Courtship displays: Acromatic contrast ($dl$); $F_{`r Population_ConspicCourtshipAC$df_num`,`r Population_ConspicCourtshipAC$df_denom`}$ = `r Population_ConspicCourtshipAC$Fstat`, p = `r p_value(Population_ConspicCourtshipAC$p_val)`; Chromatic contrast ($dS$): $F_{`r Population_ConspicCourtshipCC$df_num`,`r Population_ConspicCourtshipCC$df_denom`}$ = `r Population_ConspicCourtshipCC$Fstat`, p = `r p_value(Population_ConspicCourtshipCC$p_val)`), and this was  true across all body regions measured (See Supplemental Materials, table S1). Kenyan chameleons from populations where predation is high were significantly more crypic compared to introduced Hawaiian counterparts (Acromatic Contrast between Hawaiian and Kenyan Populations, Male-male contests: $\beta$ = `r robust_ConspicDisplayAC$beta[3]`, 95% CI = `r robust_ConspicDisplayAC_CI$CI_L[3]` to  `r robust_ConspicDisplayAC_CI$CI_U[3]`, df = `r robust_ConspicDisplayAC$df[3]`, t =  `r robust_ConspicDisplayAC$tstat[3]`, p = `r p_value(robust_ConspicDisplayAC$p_Satt[3])`, Courtship: $\beta$ = `r robust_ConspicCourtshipAC_CI$beta[3]`, 95% CI = `r robust_ConspicCourtshipAC_CI$CI_L[3]` to  `r robust_ConspicCourtshipAC_CI$CI_U[3]`, df = `r robust_ConspicCourtshipAC_CI$df[3]`, t =  `r summary(ConspicCourtshipAC)$coefficients[3, "t value"]`, p = `r p_value(robust_ConspicCourtshipAC$p_Satt[3])`, see figure 1a & b).

As a consequence of signal release, sexual signals are expected to be more easily detected by predators (i.e., snakes and birds) in Hawaii. Differential predation pressure resulting from snake and bird predation varying would also result in populations varying in their conspicuousness to different predators.  In support of these predictions, Hawaiian chameleons were significantly more conspicous to both snake and bird predators (figure 1b & c; Achromatic Contrast: ($dl$); $F_{`r Population_AC_Pred$df_num`,`r Population_AC_Pred$df_denom`}$ = `r Population_AC_Pred$Fstat`, p = `r p_value(Population_AC_Pred$p_val)`). Differences in chromatic contrast between populations, however,  depended on the predator (significant predator by population interaction: $F_{`r back_pred_CC_Pred$df_num`,`r back_pred_CC_Pred$df_denom`}$ = `r back_pred_CC_Pred$Fstat`, p = `r p_value(back_pred_CC_Pred$p_val)`). Interestingly, Kenyan and Hawaiian lizards were equally detectable by snakes, but Kenyan chameleons were significantly more cryptic from birds compared to Hawaiian chamleons (figure 1c; Snakes (Hawaii-Kenya): `r CI_hyp_test[1,2]`, 95% CI = `r CI_hyp_test[1,3]` to `r CI_hyp_test[1,4]`, p = 1; Birds (Hawaii-Kenya): `r CI_hyp_test[2,2]`, 95% CI = `r CI_hyp_test[2,3]` to `r CI_hyp_test[2,4]`, p = `r CI_hyp_test[2,8]`).

Local adaptation of sexual signals would result in Hawaiian chameleons being more conspicous to conspecifics in the Hawaiian environment compared to their native Kenyan environment. While it was not possible to conduct common garden experiments to provide definitive tests of local adaptation, our results support the local adaptation hypothesis given that Hawaiian chameleons were more conspicous to other chamleeons in the Hawaiian environment compared to the same chameleons in a Kenyan environment in both social contexts (figure 2; Contrast between Hawaiian lizard on Kenyan vs Hawaiian on Hawaiin background - Male-male contests: Achromatic ($dl$), $\beta$ = `r CI_dis_AC[1,1]`, 95% CI = `r CI_dis_AC[1,2]` to `r CI_dis_AC[1,3]`, p = `r p_displ_AC`, Chromatic ($dS$) $\beta$ = `r CI_Disp_CC[1,1]`, 95% CI = `r CI_Disp_CC[1,2]` to `r CI_Disp_CC[1,3]`, p = `r p_displ_CC`; Courtship: Achromatic ($dl$), $\beta$ = `r CI_AC_Courtship[1,1]`, 95% CI = `r CI_AC_Courtship[1,2]` to `r CI_AC_Courtship[1,3]`, p = `r p_courtship_AC`, Chromatic ($dS$) $\beta$ = `r CI_CC_Courtship[1,1]`, 95% CI = `r CI_CC_Courtship[1,2]` to `r CI_CC_Courtship[1,3]`, p = `r p_courtship_CC`).
